---
output: html_document
---

The data set looked at in this exercise was extracted from the Lahman Database using MySQL. The data includes the following Sabermetric statistics:

* PA: Plate Appearance
* OBP: On-Base Percentage
* SLG: Slugging Average
* OPS: On-Base Plus Slugging

In addition, teamID, yearID and playerID were also extracted. Only 2014 season and players with 400 PA or more were extracted to be illustrated in this exercise. 

For more information about the data, please check the [Lahman Database Documentation][id]

[id]: http://seanlahman.com/files/database/readme2014.txt

# Reading and Summarizing the Data 
```{r}
# Reading the data
setwd("/home/hduser/Sabermetrics")
OPS_2014 <- read.csv("PA.csv")
# Summary statistics of the metrics (PA, OBP, SLG, OPS)
summary(OPS_2014[, !names(OPS_2014) %in% c("yearID", "playerID", "teamID")])
# Create a dataframe of statistics using the which function to exclude playerID, yearID and teamID
OPS_2014_Stats <- OPS_2014[,-which(names(OPS_2014) %in% c("playerID", "yearID", "teamID"))]
# Check the new data frame 
head(OPS_2014_Stats)
# Again check the summary statistics
summary(OPS_2014_Stats)

# Yet a different way to keep the Sabermetric Statistics using the subset function
OPS_2014_Statistics <- subset(OPS_2014, select = c("PA", "OBP", "SLG", "OPS"))
head(OPS_2014_Statistics)
```

# Ploting the Distribution of OPS
```{r, fig.width=10, fig.height=6}
library(ggplot2)
p <- ggplot(data=OPS_2014_Stats, aes(OPS))
p + geom_histogram(aes(y = ..density..), binwidth=0.02, fill="darkblue", colour="black") + 
  geom_density(colour="red", lwd=1.2) +
  ylab(" ") +
  ggtitle("Distribution of On-Base Plus Slugging in 2014")
```
